---
title: "Personal Level screening for COPD and COPD Patient Management Survey "
format: pdf
editor: visual
editor_options: 
  chunk_output_type: console

execute:
  echo: false
  warning: false
---

# Personal Level screening for COPD

## Data pre-prepossessing

```{r}
#rm(list = ls())

```

```{r}
library(readxl)
data0<- read_excel("Copy of Personal_Level_screening_for_COPD(1).xlsx", 
    sheet = "Sheet6",skip = 1)
data0$`Educational qualification`<-factor(data0$`Educational qualification`,
                                          levels = c("No formal education",
                                                     "Secondary level",
                                                     "Undergraduate",
                                                     "Graduate",
                                                     "Postgraduate"))

#View(data0)

library(tidyverse)
library(knitr)
library(gtsummary)

#glimpse(data0)




```

### Demographic and other characteristics of participants

```{r}
##| label: tbl-tfrequency
##| tbl-cap: Demographic and other characteristics of participants

#glimpse(data0)

data0[,c(3:10,15,16,12,14)] %>% 
  tbl_summary(statistic = list(all_continuous() ~ "{mean} ({sd})"),
              label =list(`Do you have any member in your family have experienced the same condition?`="Do you have any member in your family have experienced the same condition? (Yes)" ))
```

### Summary statistics of "Duration of having bronchitis or chronic coughing with sputum from the chest in the last year" and "Duration of having bronchitis or chronic coughing with sputum from the chest "

```{r}
#| label: tbl-tsumstat
#| tbl-cap: Summary statistics of duarion of bronchitis or chronic coughing with sputum from the chest (in months)


data0 %>%select(`How many months in the last year have you had bronchitis or chronic coughing with sputum from the chest?`,`For how many years you had bronchitis or chronic coughing with sputum from the chest?`) %>% gather(key = "Variable",value = "Duration") %>% 
  group_by(Variable) %>%
  summarise(
    Min = min(Duration, na.rm = TRUE),
    Max = max(Duration, na.rm = TRUE),
    Mean = mean(Duration, na.rm = TRUE),
    SD = sd(Duration, na.rm = TRUE)
  )->s1

s1 %>% kable(digits = 2)

```

```{r}
#| label: fig-fighist
#| fig-cap: "Frequency histogram of duration of having bronchitis or chronic coughing "
#| fig-subcap: 
#|   - "Variable 1"
#|   - "Variable 2"
#| layout-ncol: 2


par(mar = c(6, 6, 4, 2))  # Adjust as needed
par(mgp = c(4, 1, 0))  # First value controls axis title distance

par(cex.lab = 1.3)  # Set larger axis labels globally
par(cex.axis=1)

hist(data0$`How many months in the last year have you had bronchitis or chronic coughing with sputum from the chest?`,breaks = 5,
     xlab = "Duration of having bronchitis or chronic coughing \n with sputum from the chest in the last year (in months)",
     main = "",col = "steelblue",ylim = c(0,59),)

hist(data0$`For how many years you had bronchitis or chronic coughing with sputum from the chest?`,breaks = 5,
     xlab = "Duration of having bronchitis or chronic \n coughing with sputum from the chest (in months) ", main = "",col = "steelblue")




```

### Do you have previous history of vaccination? if yes, write the name of vaccine

```{r}
##| label: tbl-tvaccine_stat
##| tbl-cap: Previous history of vaccination

#data0$`Do you have previous history of vaccination? if yes, write the name of vaccine.`


data0 %>% 
  separate_rows(`Do you have previous history of vaccination? if yes, write the name of vaccine.`, sep = ",") %>%  # Split comma-separated values into rows
  count(`Do you have previous history of vaccination? if yes, write the name of vaccine.`, sort = TRUE)->vaccine_stat


vaccine_stat %>% filter(`Do you have previous history of vaccination? if yes, write the name of vaccine.`=="No") %>%
  mutate(Percent=n*100/250) %>% 
  kable(col.names = c("Do you have previous history of vaccination?","N","Percent (out of 250)"),digits = 2)



  
  
```

```{r}
vaccine_stat %>% filter(`Do you have previous history of vaccination? if yes, write the name of vaccine.`!="No") %>%
  mutate(Percent=n*100/250) %>%
  kable(col.names = c("If yes, write the name of vaccine","N","Percent (out of 250)"),digits = 2)

```

```{r}
#| label: fig-fvachistry
#| fig-cap: Distribution of vaccination history
#| 
vaccine_stat %>% filter(`Do you have previous history of vaccination? if yes, write the name of vaccine.`!="No") %>%mutate(Percent=n*100/250) %>% 
  ggplot(aes(x=reorder(`Do you have previous history of vaccination? if yes, write the name of vaccine.`,Percent),y=Percent))+
  geom_col()+labs(x="Previous history of vaccination")+
  coord_flip()+
  theme_bw()+
  theme(axis.text = element_text(angle = 0,size = 12),
        axis.title = element_text(size=14))
```

### Which medication you are taking for COPD?

```{r}
#rm(list = ls())

data0<- read_excel("Copy of Personal_Level_screening_for_COPD(1).xlsx", 
    sheet = "Sheet6",skip = 1)

#glimpse(data0)
#view(data0)
#data0$`Which medication you are taking for COPD (upated)?`

data0 %>%separate_rows(`Which medication you are taking for COPD (upated)?`,sep = ",") %>% 
  count(`Which medication you are taking for COPD (upated)?`, name = "Frequency") %>% 
  #print(n=55)
  mutate(Percentage = Frequency*100/250) %>%
  #view() %>% 
  arrange(-Frequency)->medi_take
  
medi_take %>%  kable(digits = 2,col.names = c("Medication taking for COPD?`","Frequency","Percentage (out of 30)"))

```

### Corticosteroids, Oxygen therapy ,Bronchodilators data preparation

```{r}
# Data preparation for multiple response:: "`Which medication you are taking for COPD?`"| Type-Binary

# Function to one-hot encode a column
one_hot_encode <- function(data, column_name) {
  data %>%
    separate_rows(!!sym(column_name), sep = ",") %>%  # Split into multiple rows
    mutate(Value = 1) %>%
    pivot_wider(names_from = !!sym(column_name), values_from = Value, values_fill = 0)  # Convert to wide format
}


###########################
#data0$`Which medication you are taking for COPD?`
library(janitor)
clean_names(data0)->data1

#data1$which_medication_you_are_taking_for_copd
#data1$do_you_have_a_previous_history_of_smoking

# One-hot encode Symptoms
df_medication <- one_hot_encode(data1,"which_medication_you_are_taking_for_copd")

df_encoded <- data1 %>% select(id, age,gender,do_you_have_a_previous_history_of_smoking) %>%
  left_join(df_medication %>% select(-age,-gender,-do_you_have_a_previous_history_of_smoking), by = "id")  # Remove duplicate columns

#df_encoded %>% view()


#df_encoded$`Oxygen therapy`

df_encoded %>% select(age,gender,do_you_have_a_previous_history_of_smoking,Corticosteroids,`Oxygen therapy`,Bronchodialators)->association_data




```

### A) Is there any association between medication (Corticosteroids) and smoking?

```{r}
#| label: tbl-tasso_corti
#| tbl-cap: Association between smoking status and prevelance of Corticosteroids

# Function to extract chi-square test statistic
chisq_test_stat <- function(data, variable, by, ...) {
  chisq.test(table(data[[variable]], data[[by]])) %>% 
  broom::tidy() %>%
    select(statistic, p.value)
}


association_data %>% select(Corticosteroids,do_you_have_a_previous_history_of_smoking) %>%
 # mutate_if(is.numeric,as.factor) %>% 
  tbl_summary(by=do_you_have_a_previous_history_of_smoking,
              label=list(do_you_have_a_previous_history_of_smoking~"Smoking status",
                         Corticosteroids~"Corticosteroids (Yes)"))%>%
  add_stat(fns = everything() ~ chisq_test_stat) %>% 
  modify_header(statistic = "**Chi-square**", p.value = "**p-value**") %>%
  add_p(pvalue_fun = ~ style_pvalue(.x, digits = 3)) %>% 
  modify_spanning_header(starts_with("stat_") ~ "**Smoking status**")  # Column variable name



```

The prevalence of *Corticosteroids* was the most in group of participants/ patients who previously smoke (41%), followed by the participants who still smoking (31%) and who did not smoke (28%). But the proportions of prevalence of *Corticosteroids* are not significantly different among the groups, $\chi^2=3.68$, p-value=0.159 (@tbl-tasso_corti).

### B) Is there any association between medication (Oxygen therapy) and smoking?

```{r}
association_data %>% select(`Oxygen therapy`,do_you_have_a_previous_history_of_smoking) %>% 
  tbl_summary(by=do_you_have_a_previous_history_of_smoking,
              label=list(do_you_have_a_previous_history_of_smoking~"Smoking status",
                         `Oxygen therapy`~"Oxygen therapy (Yes)"))%>%
  add_stat(fns = everything() ~ chisq_test_stat) %>% 
  modify_header(statistic = "**Chi-square**", p.value = "**p-value**") %>%
  add_p(pvalue_fun = ~ style_pvalue(.x, digits = 3)) %>%
  modify_spanning_header(starts_with("stat_") ~ "**Smoking status**")  # Column variable name


```

*Please describe same as @tbl-tasso_corti.*

### C) Is there any association between medication (Bronchodialators) and smoking?

```{r}

association_data %>% 
  select(Bronchodialators,do_you_have_a_previous_history_of_smoking) %>% 
  tbl_summary(by=do_you_have_a_previous_history_of_smoking,
              label=list(do_you_have_a_previous_history_of_smoking~"Smoking status",
                         Bronchodialators~"Bronchodilators (Yes)"))%>%
  add_stat(fns = everything() ~ chisq_test_stat) %>% 
  modify_header(statistic = "**Chi-square**", p.value = "**p-value**") %>%
  add_p(pvalue_fun = ~ style_pvalue(.x, digits = 3)) %>%
  modify_spanning_header(starts_with("stat_") ~ "**Smoking status**")  # Column variable name
```

```{r}

glm(Corticosteroids~age+gender+do_you_have_a_previous_history_of_smoking,
    family="binomial", data=association_data)->glm1
#tbl_regression(glm1, exponentiate = TRUE)


glm(`Oxygen therapy`~age+gender+do_you_have_a_previous_history_of_smoking,
    family="binomial", data=association_data)->glm2
#tbl_regression(glm2,exponentiate = TRUE)

```

# COPD Patient Management Survey_old (n=30)


```{r}
#rm(list = ls())

```

### Specialization

Specialization of physicians.

```{r}
library(readxl)

Survey <- read_excel("COPD Patient Management Survey.xlsx", 
    sheet = "Doctor")

#glimpse(Survey)

Survey %>% 
  separate_rows(Specialization,sep = ",") %>% 
  count(Specialization, name = "Frequency") %>% # Count occurrences
  mutate(Percentage = Frequency*100/30) %>%arrange(-Frequency) %>% 
  kable(digits = 2,col.names = c("Specialization","Frequency","Percentage (out of 30)"))


```

### The most common age group of COPD patient, Age group comes for recurrent treatment

```{r}
Survey[,5:6] %>% tbl_summary()
```

### What are the most common symptoms presented by COPD patients?

```{r}


Survey %>% 
  separate_rows(`What are the most common symptoms presented by COPD patients?`,sep = ";") %>% 
  count(`What are the most common symptoms presented by COPD patients?`, name = "Frequency") %>% # Count occurrences
  mutate(Percentage = Frequency*100/30) %>%arrange(-Frequency) %>% 
  kable(digits = 2,col.names = c("Most common symptoms presented by COPD patients","Frequency","Percentage (out of 30)"))

```

### Write symptom if the previous answer is "others"

```{r}

Survey %>% 
  separate_rows(`write symptom if the previous answer is "others"`,sep = ",") %>% 
  count(`write symptom if the previous answer is "others"`, name = "Frequency") %>% # Count occurrences
  mutate(Percentage = Frequency*100/30) %>%arrange(-Frequency) %>% 
  kable(digits = 2,col.names = c("Symptom if the previous answer is others","Frequency","Percentage (out of 30)"))
```

### Which diagnostic tests do you commonly use for diagnosis and monitoring COPD?

```{r}
#glimpse(Survey)

Survey %>% 
  separate_rows(`Which diagnostic tests do you commonly use for diagnosis and monitoring COPD?`,sep = ";") %>% 
  count(`Which diagnostic tests do you commonly use for diagnosis and monitoring COPD?`, name = "Frequency") %>% # Count occurrences
  mutate(Percentage = Frequency*100/30) %>%arrange(-Frequency) %>% 
  kable(digits = 2,col.names = c("Diagnostic tests do you commonly use for diagnosis and monitoring COPD","Frequency","Percentage (out of 30)"))
```

### write diagnostic test if the previous answer is "others"

```{r}

#glimpse(Survey)

Survey %>% 
  separate_rows(`write diagnostic test if the previous answer is "others"`,sep = ", ") %>% 
  count(`write diagnostic test if the previous answer is "others"`, name = "Frequency") %>% # Count occurrences
  mutate(Percentage = Frequency*100/30) %>%arrange(-Frequency) %>% 
  kable(digits = 2,col.names = c("Diagnostic test if the previous answer is others","Frequency","Percentage (out of 30)"))

```

### Which medication do you typically prescribe for COPD patients?

```{r}
#glimpse(Survey)

Survey %>% 
  separate_rows(`Which medication do you typically prescribe for COPD patients?`,sep = ";") %>% 
  count(`Which medication do you typically prescribe for COPD patients?`, name = "Frequency") %>% # Count occurrences
  mutate(Percentage = Frequency*100/30) %>%arrange(-Frequency) %>% 
  kable(digits = 2,col.names = c("Medication do you typically prescribe for COPD patients?","Frequency","Percentage (out of 30)"))
```

### If previous answer is other, Kindly specify-

Two answers-Long acting and Long-acting beta-agonists.

### What is the average cost of a month's worth of COPD medication for a patient (estimate in local currency)?

```{r}
Survey %>% select(`What is the average cost of a month's worth of COPD medication for a patient (estimate in local currency)?`) %>% 
  tbl_summary()
```

### Do you observe any differences in medication response between smokers and non-smokers? If yes please describe-

```{r}
Survey %>% select(`Do you observe any differences in medication response between smokers and non-smokers?`, `If Yes, Please describe-`) %>% 
  tbl_summary()
```

### How often do you recommend follow up tests for COPD patient (e.g. spirometry)?+

### How frequently do COPD patients require hospitalization due to exacerbations? +

### What lifestyle changes do you most commonly recommend to COPD patients? +

### Please specify if previous answer is "Others" +

### Are there any new treatments or medication for COPD that you find promising?

```{r}
Survey %>% select(`How often do you recommend follow up tests for COPD patient (e.g. spirometry)?`,`How frequently do COPD patients require hospitalization due to exacerbations?`,`What lifestyle changes do you most commonly recommend to COPD patients?`,`Please specify if previous answer is "Others"`,`Are there any new treatments or medication for COPD that you find promising?`) %>% 
  tbl_summary()

```

# COPD Patient Management Survey_updated (n=30)


```{r}
#rm(list = ls())

```

### Specialization

Specialization of physicians.

```{r}
library(readxl)

Survey2 <- read_excel("COPD Patient Management doctors Survey.xlsx", 
    sheet = "updated",skip = 1)

#glimpse(Survey)

Survey2 %>% 
  separate_rows(Specialization,sep = ",") %>% 
  count(Specialization, name = "Frequency") %>% # Count occurrences
  mutate(Percentage = Frequency*100/30) %>%arrange(-Frequency) %>% 
  kable(digits = 2,col.names = c("Specialization","Frequency","Percentage (out of 30)"))


```

### The most common age group of COPD patient, Age group comes for recurrent treatment

```{r}
Survey2[,3:4] %>% tbl_summary()
```

### What are the most common symptoms presented by COPD patients?

```{r}

Survey2 %>% 
  separate_rows(`What are the most common symptoms presented by COPD patients?`,sep = ";") %>% 
  count(`What are the most common symptoms presented by COPD patients?`, name = "Frequency") %>% # Count occurrences
  mutate(Percentage = Frequency*100/30) %>%arrange(-Frequency) %>% 
  kable(digits = 2,col.names = c("Most common symptoms presented by COPD patients","Frequency","Percentage (out of 30)"))

```

### Write symptom if the previous answer is "others"

```{r}

Survey2 %>% 
  separate_rows(`write symptom if the previous answer is "others"`,sep = ",") %>% 
  count(`write symptom if the previous answer is "others"`, name = "Frequency") %>% # Count occurrences
  mutate(Percentage = Frequency*100/30) %>%arrange(Frequency) %>% 
  kable(digits = 2,col.names = c("Symptom if the previous answer is others","Frequency","Percentage (out of 30)"))
```

### Which diagnostic tests do you commonly use for diagnosis and monitoring COPD?

```{r}
#glimpse(Survey)

Survey2 %>% 
  separate_rows(`Which diagnostic tests do you commonly use for diagnosis and monitoring COPD?`,sep = ";") %>% 
  count(`Which diagnostic tests do you commonly use for diagnosis and monitoring COPD?`, name = "Frequency") %>% # Count occurrences
  mutate(Percentage = Frequency*100/30) %>%arrange(-Frequency) %>% 
  kable(digits = 2,col.names = c("Diagnostic tests do you commonly use for diagnosis and monitoring COPD","Frequency","Percentage (out of 30)"))
```

### write diagnostic test if the previous answer is "others"

```{r}

#glimpse(Survey)

Survey2 %>% 
  separate_rows(`write diagnostic test if the previous answer is "others"`,sep = ", ") %>% 
  count(`write diagnostic test if the previous answer is "others"`, name = "Frequency") %>% # Count occurrences
  mutate(Percentage = Frequency*100/30) %>%arrange(-Frequency) %>% 
  kable(digits = 2,col.names = c("Diagnostic test if the previous answer is others","Frequency","Percentage (out of 30)"))

```

### Which medication do you typically prescribe for COPD patients?

```{r}
#glimpse(Survey)

Survey2 %>% 
  separate_rows(`Which medication do you typically prescribe for COPD patients?`,sep = ";") %>% 
  count(`Which medication do you typically prescribe for COPD patients?`, name = "Frequency") %>% # Count occurrences
  mutate(Percentage = Frequency*100/30) %>%arrange(-Frequency) %>% 
  kable(digits = 2,col.names = c("Medication do you typically prescribe for COPD patients?","Frequency","Percentage (out of 30)"))
```

### If previous answer is other, Kindly specify-

Two answers-Long acting and Long-acting beta-agonists.

### What is the average cost of a month's worth of COPD medication for a patient (estimate in local currency)?

```{r}
Survey2$`What is the average cost of a month's worth of COPD medication for a patient (estimate in local currency)?`<-factor(Survey2$`What is the average cost of a month's worth of COPD medication for a patient (estimate in local currency)?`,
                                                                                                                            levels = c("About 5000tk","5000-10000 tk",
                                                                                                                                       "10000-15000tk","15000-20000 tk",
                                                                                                                                       "Depends on treatment plan"))

Survey2 %>% select(`What is the average cost of a month's worth of COPD medication for a patient (estimate in local currency)?`) %>% 
  tbl_summary()
```

### Do you observe any differences in medication response between smokers and non-smokers? If yes please describe-

```{r}
Survey2 %>% select(`Do you observe any differences in medication response between smokers and non-smokers?`, `If Yes, Please describe-`) %>% 
  tbl_summary()
```

### How often do you recommend follow up tests for COPD patient (e.g. spirometry)?+

### How frequently do COPD patients require hospitalization due to exacerbations? +

### What lifestyle changes do you most commonly recommend to COPD patients? +

### Please specify if previous answer is "Others" +

### Are there any new treatments or medication for COPD that you find promising?

```{r}
Survey2 %>% select(`How often do you recommend follow up tests for COPD patient (e.g. spirometry)?`,`How frequently do COPD patients require hospitalization due to exacerbations?`,`What lifestyle changes do you most commonly recommend to COPD patients?`,`Please specify if previous answer is "Others"`,`Are there any new treatments or medication for COPD that you find promising?`) %>% 
  tbl_summary()

```


