---
title: "Central Limit Theorem Simulation"
format: 
  html: 
    page-layout: full
filters:
  - shiny
engine: knitr
---

```{r}
library(shiny)
library(ggplot2)
```


```{r}
ui <- fluidPage(
  titlePanel("Central Limit Theorem Simulation"),
  sidebarLayout(
    sidebarPanel(
      selectInput("dist", "Choose Distribution:",
                  choices = c("Uniform" = "unif", "Exponential" = "exp", "Poisson" = "pois")),
      sliderInput("n", "Sample Size (n):", min = 5, max = 100, value = 30),
      sliderInput("samples", "Number of Samples:", min = 100, max = 5000, value = 1000),
      actionButton("simulate", "Simulate")
    ),
    mainPanel(
      plotOutput("histPlot")
    )
  )
)

server <- function(input, output) {
  observeEvent(input$simulate, {
    output$histPlot <- renderPlot({
      n <- input$n
      samples <- input$samples
      dist <- input$dist
      
      sample_means <- numeric(samples)
      
      for (i in 1:samples) {
        if (dist == "unif") {
          sample_means[i] <- mean(runif(n))
        } else if (dist == "exp") {
          sample_means[i] <- mean(rexp(n, rate = 1))
        } else if (dist == "pois") {
          sample_means[i] <- mean(rpois(n, lambda = 4))
        }
      }
      
      ggplot(data.frame(sample_means), aes(x = sample_means)) +
        geom_histogram(aes(y = ..density..), bins = 30, fill = "skyblue", color = "white") +
        stat_function(fun = dnorm, args = list(mean = mean(sample_means), sd = sd(sample_means)),
                      col = "red", size = 1.2) +
        labs(title = "Sampling Distribution of the Mean", x = "Sample Means", y = "Density") +
        theme_minimal()
    })
  })
}

shinyApp(ui = ui, server = server)